---
title: "Hog Production (Fable)"
author: "ODH"
format: pdf
execute:
  echo: false
  warning: false
editor: visual
---

```{r}
library(fable)
library(tidyverse)
library(tsibble)
library(feasts)
library(forecast)
```

```{r}
hp <- read_csv("exploring-data/hp.csv")
hp <- tibble(hp = hp$HP, date = seq(from = as.Date("2013/1/1"), by = "quarter", length.out = 40))
hp <- hp |>
  mutate(quarter = yearquarter(date)) |>
  select(-date) |>
  as_tsibble(index = quarter)
hp_ts <- ts(hp$hp, start = 2013, frequency = 4)
```

# Hog Production Time Series

```{r}
autoplot(hp, hp)
```

# Seasonal Plot

```{r}
hp |> gg_season(hp, period = "year") 
```

# STL Decomposition

```{r}
hp |>
  model(
    STL(hp ~ trend() +
          season(),
    robust = TRUE)) |>
  components() |>
  autoplot()
```

# Test of Stationarity

## The Data

```{r}
hp |>
  features(hp, unitroot_kpss)
```

The pvalue is less than 0.05. The data is not stationary. The data needs to be differenced.

## 1st Order Differencing

```{r}
hp |>
  features(difference(hp), unitroot_kpss)
```

The pvalue is less than 0.05. The data is not stationary. The data needs to be differenced.

## 2nd Order Differencing

```{r}
hp |>
  features(difference(difference(hp)), unitroot_kpss)
```

The pvalue is greater than 0.05. The data is stationary.

## ACF and PACF

```{r}
hp |>
  gg_tsdisplay(difference(difference(hp)), plot_type='partial', lag = 24)
```

ACF suggests an ARIMA(0,2,3) and PACF suggests an ARIMA(4,2,0).

# Models

```{r}
hp_fit <- hp |>
  model(arima023 = ARIMA(hp ~ pdq(0,2,3)),
        arima420 = ARIMA(hp ~ pdq(4,2,0)),
        arima021 = ARIMA(hp ~ pdq(0,2,1)),
        arima120 = ARIMA(hp ~ pdq(1,2,0)),
        arima121 = ARIMA(hp ~ pdq(1,2,1)),
        arima321 = ARIMA(hp ~ pdq(3,2,1)),
        arima021100 = ARIMA(hp ~ pdq(0,2,1) + PDQ(1,0,0)),
        arima121100 = ARIMA(hp ~ pdq(1,2,1) + PDQ(1,0,0)),
        stepwise = ARIMA(hp),
        search = ARIMA(hp, stepwise=FALSE),
        search1 = ARIMA(hp, stepwise=FALSE, approximation = FALSE))
```

```{r}
fit1 <- Arima(hp_ts, order=c(0,2,1))
fit1
```

```{r}
autoplot(fit1)
```

```{r}
glance(hp_fit) |>
  arrange(AICc) |>
  select(.model:BIC)
```

The stepwise and search models are autoarima models. They are actually the same model. The model is shown below.

```{r}
hp_fit |>
  select(arima021)
```

```{r}
gg_arma(hp_fit |>
          select(arima021))
```

```{r}
hp_fit |>
  select(arima121) |>
  report()
```

The best model is the autoarima model with the least AIC.

# Residual Checking

## arima023

```{r}
hp_fit |>
  select(arima023) |>
  gg_tsresiduals()
```

ACF are within the threshold limits, indicating that residuals are behaving like white noise.

```{r}
augment(hp_fit) |>
  filter(.model=='arima023') |>
  features(.innov, ljung_box)
```

The pvalue is greater than 0.05, suggesting that the residuals are white noise.

## arima420

```{r}
hp_fit |>
  select(arima420) |>
  gg_tsresiduals()
```

ACF are within the threshold limits, indicating that residuals are behaving like white noise.

```{r}
augment(hp_fit) |>
  filter(.model=='arima420') |>
  features(.innov, ljung_box)
```

The pvalue is greater than 0.05, suggesting that the residuals are white noise.

## AutoARIMA

```{r}
hp_fit |>
  select(search) |>
  gg_tsresiduals()
```

ACF are within the threshold limits, indicating that residuals are behaving like white noise.

```{r}
augment(hp_fit) |>
  filter(.model=='search') |>
  features(.innov, ljung_box)
```

The pvalue is greater than 0.05, suggesting that the residuals are white noise.

# 5 Quarters Forecast

## AutoARIMA Forecast

```{r}
hp_fit |>
  forecast(h=5) |>
  filter(.model=='search') |>
  autoplot(hp)
```

## ARIMA(4,2,0) Forecast

```{r}
hp_fit |>
  forecast(h=5) |>
  filter(.model=='arima420') |>
  autoplot(hp)
```

## ARIMA(0,2,3) Forecast

```{r}
hp_fit |>
  forecast(h=5) |>
  filter(.model=='arima023') |>
  autoplot(hp)
```
